Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-01-29 unstructured.tex
2013.04.03 (Alex) More on the unstructured algorithms text.

In this section, we introduced the topology mismatching problem from the
perspective of unstructured P2P systems and we visited a series
of algorithms that have been proposed in the literature to tackle its negative
influence on performance. In surveying the above algorithms, we identify
four main methodologies that are used either selectively or in combination
by each algorithm.  
These methodologies are:
\begin{enumerate}[\itshape i\upshape)]
  \item topology adaptation
  \item forwarding optimization
  \item caching and replication, and
  \item landmarking
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Topology adaptation}
\begin{figure}[ht]
\centering
\subfigure[Inefficient overlay topology.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
  \label{figure:topology-adaptation:before}
}\qquad\qquad
\subfigure[Efficient overlay topology after adaptation.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
  \label{figure:topology-adaptation:after}
}
\caption{A simple example of how an overlay topology can be adapted.}
\label{figure:topology-adaptation}
\end{figure}

The topology adaptation based protocols modify the topology of the P2P network
using various techniques. The two most commonly used approaches to 
topology adaptation are creating
\emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
physically close nodes.

Narada and subsequent algorithms like AOTO, LTM, SBO, try to solve the problem
by building a richer connected graph and forming minimum spanning trees over
this graph that can efficiently route messages among peers. Even though spanning
trees provide efficient query performance, their maintenance costs have 
proven to be a significant limitation. This is why the original Narada 
algorithm 
was intended for small
groups and is thus, not scalable.  Algorithms like AOTO, LTM or SBO try to
overcome these obstacles with clever tricks like forming minimum spanning trees
for the two-hop neighbors ($N^2$) of each node, partitioning the graph into two
random groups where each group is responsible for different tasks, or performing
local optimizations dynamically on the overlay graph. The advantage of building
minimum spanning trees is that they maintain the connectivity on the network in
an efficient manner while still preserving the overall search scope. However,
their construction and the update costs, especially in dynamic, high-churn
environments, cause large additional traffic overhead on the underlying
network \cite{CRZ2000,CRSZ2001,CRSZ2002}.

The cluster-based approaches, on the other hand, link physically
closer nodes to each other. T2MC, for example, uses trace-route logs and DDNO
uses domain names to cluster nearby nodes. Unfortunately, commonly used
methods for proximity detection across Internet do not always return
reliable results and therefore mapping accuracy is not guaranteed. Moreover,
trace-route 
is a heavy-weight utility and many network equipment vendors do not allow
trace-route calls at all. The most problematic aspect of clustering, though, is
its nature per se. Limited connectivity among the various local domains can
significantly shrink the search scope, negatively affecting the query response
time that the P2P end-user experiences. Techniques, like DDNO, that try to 
balance the efficiency of the clustering approach with improved
connectivity have also been proposed.  Specifically, DDNO attempt to address this
limitation by forcing half of each node's connections to be with other, randomly
selected, nodes over the network.

In brief, topology adaptation ultimately aims to reduce the 
average cost of the paths from one node to another so that 
the overlay network is better fit on top of the underlying IP network.
This is fair in the general case but not always. For example an
adaptive algorithm can exchange a slow edge in the overlay with a faster one
thus reducing the average latency of the network communication. The pitfall here
is that this new virtual link can traverse a fast AS-to-AS
link meaning that even though message
round-trip-time is reduced, it has more cost in terms of communication network economics
and management.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Forwarding optimization}

\begin{figure}[ht]
\centering
\subfigure[Typical blind flooding.] {
\includegraphics[scale=0.25]{img/pdf/forwarding-optimization-before.pdf}
  \label{figure:forwarding-optimisation:before}
}\qquad\qquad
\subfigure[A node can decide where to forward the messages.] {
  \includegraphics[scale=0.25]{img/pdf/forwarding-optimization-after.pdf}
  \label{figure:forwarding-optimisation:after}
}
\caption{Forwarding optimization in constraining message flooding.}
\label{figure:forwarding-optimisation}
\end{figure}

% Researchers initially started considering alternatives for services who
% implement one-to-many communication schemes, like IP multi-cast, on the
% application layer (i.e. Narada). Application level multi-cast protocols
% are generic protocols that can be applied to both P2P file sharing systems,
% as well as content distribution networks. Application layer multi-cast,
however,
% is not as efficient as the one enabled by IP and still struggles to ease the
% difficulties imposed by the topology mismatch problem.

@@

Forwarding-optimization-based
approaches, propose intelligent forwarding link selection for a message's next
hop across the routing path. The selection criteria vary depending on the
algorithm, but they mainly use one or more statistical heuristics. These can
range from simple obvious ones, like the latency of the link between the nodes or
the exposure of specific features, such as high processing bandwidth, high
capacity, or low latency, to more sophisticated alternatives that may even take
into account application level requirements such as the reliability or
friendship. Figure~\ref{figure:forwarding-optimisation} shows a simple example
of how forwarding optimization can work.
Sub-figure~\ref{figure:forwarding-optimisation:before} shows an algorithm that
floods the entire network for search of an object the existence of which is
denoted by black nodes. Alternatively, a node can decide to forward its
messages not to every output link but to a specific or subset of its output
links. In Sub-figure~\ref{figure:forwarding-optimisation:after}, the node in the
middle, does not flood its neighbors. Instead it picks one of them. The dashed
line on the right, shows that, for this routing process, the algorithm has
rendered two potential forwarding paths as inefficient (e.g., the target nodes
show overloading signs). These approaches have the advantage of enhancing the
search responsiveness and reducing the aggregate resource usage of the physical
network. On the other hand, they suffer from drastic reduction of the search
scope (in Sub-figure~\ref{figure:forwarding-optimisation:after} the object on
the far right is not reached) thus limiting the scalability of the whole
network. Moreover they do not actually address the problem of topology mismatch
since they do not give any guarantees that overlay and underlying topologies are
aligned with each other let alone quantify the mismatch and try to alleviate.
Like all other approaches, forwarding based methods are commonly applied in
conjunction with other methods to improve the quality of the P2P systems.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Caching and replication}

\begin{figure}[ht]
\centering
\subfigure[Indexing can reduce the cost of the last hop.] {
\includegraphics[scale=0.3]{img/pdf/replication-index.pdf}
  \label{figure:replication:index}
}\qquad\qquad
\subfigure[Data replication along the path of a successful query.] {
  \includegraphics[scale=0.3]{img/pdf/replication-data.pdf}
  \label{figure:replication:data}
}
\caption{Index and data replication strategies.}
\label{figure:replication}
\end{figure}

Caching is a widely used technique to exploit locality and minimize redundant
transfer of data, previously successfully adopted by web and file server
application environments. Since peers in a P2P system also operate as servers,
it is intuitively expected that P2P file sharing systems can also benefit from
caching in improving performance and reducing overall resource usage. However,
the design of caches in this context is non-trivial compared to the web-based
caching. Due to the fact that each node can act as both a server and a client,
two important issues have to be considered at design time. First, the lifetime
of a query is short, as the nodes join and leave frequently. Second, the result
of a single query string is not always the same, as it is dependent on the
source of the query, the TTL value set for the messages, the current
interconnection of peers and the high volatility of the environment. Thus,
to develop a successful caching system for a P2P architecture, these
parameters also have to be considered. P2P caching/replication can be applied
at two different levels, namely caching indices or pointers to data (see
Sub-figure~\ref{figure:replication:index}) or caching the data itself (see
Sub-figure~\ref{figure:replication:data}). Successful implementations have
already been developed in some commercial P2P systems, like KazaA, which uses
caching of indices on its super-peer level. Unfortunately, even though the
state of the art in P2P algorithms using caching methods reduce the resource
usage of the network, they also fail to consistently address typical problems
of a topologically mismatched environment such as duplication of messages.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/landmarking.pdf}
\caption{Landmark binning during node bootstrap.}
\label{figure:landmarking}
\end{figure}

In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
distance measurement method to position themselves with respect to ``a priori''
known servers (the landmark servers) on the Internet. Landmark servers are used
by nodes to estimate their positions based on the intuitive assumption that
nodes
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network. In Figure~\ref{figure:landmarking} the newly
arriving peer, measures its distance to an array of landmark point denoted as
$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
order to choose the already participating peers with which to connect to,
compares its list to the list of the potential neighbor. Peers with similar
measured distances to the these landmark points are likely to be close to
each other as well.

Landmark based protocols have four important drawbacks. First, the network
delay is not a reliable distance estimation method. For example, based on the
load on the network, the delay to certain nodes or networks can change from time
to time, which will eventually affect the distance measurements and wrong
measurements will lead to wrong estimated positions for the nodes, or incorrect
and non optimal clusterings of the nodes. Second, relying on predefined nodes
make the whole paradigm not fully distributed and the landmark system prone to
become a single point of failure. Third, using
landmark servers requires costly installation and maintenance of landmark
infrastructure across the whole Internet and for all the different autonomous
system domains. As popular P2P file sharing applications usually have millions
of peers connected at any time, the network costs
of maintaining these landmark servers will likely be quite high. A possible solution to
the scalability problem of the static landmark servers is to use ordinary nodes
as dynamic landmarks once they estimate their own positions. Even though this
approach scales much better than static landmark servers, the measurement
accuracy problem affects the overall performance of the system. Moreover,it can
be characterized as a rather coarse-grained approximation, therefore not
particularly well suited for detecting the correct positions of nodes within
close distance to each other.

>>>>

While surveying efforts to overcome the mismatch problem
in the area of unstructured \p\ systems,
we came to identify four key methodologies 
utilized by the discussed approaches; these methodologies are:
%%AD shorten below as above...
% In this section, we introduced the topology mismatching problem from the
% perspective of unstructured P2P systems and we visited a series
% of algorithms that have been proposed in the literature to tackle its negative
% influence on performance. In surveying the above algorithms, we identify
% four main methodologies that are used either selectively or in combination
% by each algorithm.  
% These methodologies are:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{enumerate}[\itshape i\upshape)]
  \item topology adaptation
  \item forwarding optimization
  \item caching and replication, and
  \item landmarking
\end{enumerate}
Each surveyed approach uses either one or more of the
aforementioned four methodologies. 
Below, we discuss how the proposed protocols use 
elements of the four methodologies and we offer 
a summary qualitative comparison for all surveyed approaches 
applicable for \emph{unstructured} \p \emph{systems}.
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Topology Adaptation Methodology}

Protocols based on topology adaptation 
modify the topology of the \p\ network
using various techniques. 
The two most commonly used approaches to 
topology adaptation are creating
\emph{spanning tree}s using connection graphs 
and creating \emph{clusters} of physically close nodes.

\emph{Narada} and subsequent algorithms including 
\emph{AOTO, LTM, SBO}, try to solve the problem
by building a richer connected graph and forming 
minimum spanning trees over
this graph that can efficiently route messages among peers. 
Even though spanning trees provide efficient query performance, 
their maintenance costs have proven to be a significant limitation. 
This is why the original \emph{Narada} proposal
focused on small groups and thus, is not scalable.  
%%
The \emph{AOTO, LTM} and \emph{SBO} try to 
% Algorithms like AOTO, LTM or SBO try to %%AD this "like" is colloquial 
					  %%   try to avoid it if you can!
overcome these obstacles with 
ingeneous schemes
% clever tricks %%AD this is a bit colloquial
that entail forming minimum spanning trees
for the $2$-hop neighbors ($N^2$) for each node, 
%%AD u imply each has N neighbors above right???
partitioning the graph into two
random groups where each group is responsible for different tasks, 
or performing local optimizations dynamically on the overlay graph. 
%%AD what are these "optimizations" (in 2-3 words  - in sentence)
The advantage of building minimum spanning trees is that 
they maintain the connectivity on the network in
an efficient manner while still preserving the overall search scope. 
However, their construction and the update costs, 
especially in dynamic, high-churn environments, 
causes large additional traffic overhead on the underlying
network~\cite{CRZ2000,CRSZ2001,CRSZ2002}.

%%
%%AD this figure HAS to be referenced in the text!!! 
%%   ALSO - is this ANY different from Fig 4 of the paper??? 
%%   WHY this Fig is being repeated here??? Grand Mystery!!!!
%%	Perhaps we should put one fig similar but more sexy??? ;-)
\begin{figure}[ht]
\centering
\subfigure[Inefficient overlay topology.] {
  \includegraphics[scale=0.3]{img/pdf/topology-adaptation-before.pdf}
  \label{figure:topology-adaptation:before}
}\qquad\qquad
\subfigure[Efficient overlay topology after adaptation.] {
  \includegraphics[scale=0.3]{img/pdf/topology-adaptation-after.pdf}
  \label{figure:topology-adaptation:after}
}
\caption{A simplified example on how overlay topology adaptation may improve matters.}
% A simple example of how an overlay topology can be adapted.}
\label{figure:topology-adaptation}
\end{figure}
%%

The cluster-based approaches, on the other hand, 
link physically-close nodes to each other. 
\emph{T2MC}, for example, uses trace-route logs and 
\emph{DDNO} exploits domain names to cluster nodes in proximity. 
Unfortunately, commonly used
methods for proximity detection across Internet do not always return
reliable results and therefore, mapping accuracy is not guaranteed. 
Moreover, trace-route is a computing resource ``heavy-weigth'' utility 
and so  a number of network equipment vendors do not allow
trace-route calls at all. %%AD do you have a refeerence or two in here???
%%
The most problematic aspect of clustering, though, is
its nature per se. 
Limited connectivity among the various local domains can
significantly shrink the search scope, negatively affecting the query response
time that the \p\ user experiences. 
% Techniques, like DDNO, that try to 
Among others, \emph{DDNO} tries to 
balance the efficiency of the clustering approach 
with enhanced node connectivity. 
In particular, \emph{DDNO} attemps to address this limitation
by forcing half of each node's connections to be with other, randomly
selected, nodes in the network.
%%AD you mean the "logical network"? the p2p network???

As a methodology, 
topology adaptation ultimately aims to reduce the 
average path traversing cost from one node to another.
In doing so, topology adaptation  re-arranges the 
overlay network so that it becomes a better fit 
for the underlying \emph{IP} network
as Figure~\ref{figure:topology-adaptation} depicts.
Despite the fact that the above is a plausible proposition,
it is not always advantageous.
% This is fair in the general case but not always. 
%%
For example, an topology adaptive algorithm can exchange 
a slow edge in the overlay with a faster one
thus reducing the average latency of the network communication. 
The pitfall here is that this new virtual link 
can traverse a fast \emph{AS}--to--\emph{AS} 	
			%%AD you mean Autonomous Systems here?? 
		  	%%    What is this AS???
link meaning that even though message
round-trip-time is reduced, it has more cost 
in terms of communication network economics and management.
%%AD just check the above sentence for correctness..

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Forwarding Optimization Methodology}

% Researchers initially started considering alternatives for services who
% implement one-to-many communication schemes, like IP multi-cast, on the
% application layer (i.e. Narada). Application level multi-cast protocols
% are generic protocols that can be applied to both P2P file sharing systems,
% as well as content distribution networks. Application layer multi-cast,
% however,
% is not as efficient as the one enabled by IP and still struggles to ease the
% difficulties imposed by the topology mismatch problem.

@@

%%AD I think you will need to sprinkle some references of those surveyed 
%%	in this subsubsection.. As opposed to what done above, 
%%	this subsubsection has no reference in the text!!!!
Approaches based on forwarding-optimization 
propose intelligent forwarding link selection for a message's next
hop across a routing path. 
The selection criteria vary depending on the
algorithm, but they mainly use one or more statistical heuristics. 
These can range from simple and rather obvious ones, 
including the latency of the link between the nodes or
the exposure of specific features such as high processing bandwidth, high
capacity, or low latency, 
to more sophisticated alternatives that may even take
into account application level requirements such as 
the reliability or friendship.

Figure~\ref{figure:forwarding-optimisation} shows a simple example
of how forwarding optimization may work.
%%
\begin{figure}[ht]
\centering
\subfigure[Typical blind flooding.] {
\includegraphics[scale=0.20]{img/pdf/forwarding-optimization-before.pdf}
  \label{figure:forwarding-optimisation:before}
}\qquad\qquad
\subfigure[A node can decide where to forward the messages.] {
  \includegraphics[scale=0.20]{img/pdf/forwarding-optimization-after.pdf}
  \label{figure:forwarding-optimisation:after}
}
\caption{Forwarding optimization in constraining message flooding.}
\label{figure:forwarding-optimisation}
\end{figure}
%%
Fig.~\ref{figure:forwarding-optimisation:before} 
% Sub-figure~\ref{figure:forwarding-optimisation:before} 
shows a protocol that simply floods the entire network 
in search of object found in the nodes colored in black.
% for search of an object the existence of which is
% denoted by black nodes. 
Alternatively, a node can decide to forward its
messages not to every output link but to a specific 
or subset of its outgoing links. 
In Sub-figure~\ref{figure:forwarding-optimisation:after}, the node in the
In Fig.~\ref{figure:forwarding-optimisation:after}, 
the node in the middle, does not flood its neighbors as instead 
in only picks one of them.
The dashed line on the right, shows that, 
for this routing process, the protocol has
rendered two potential forwarding paths as inefficient (e.g., the target nodes
show overloading signs). 
On one hand, forward
optimisation approaches have 
the advantage of enhancing the
search responsiveness and reducing the aggregate resource 
usage of the physical network. On the other, 
they suffer from drastic reduction of the search
scope (in Fig.~\ref{figure:forwarding-optimisation:after} the object on
the far right is not reached) thus limiting the scalability of the whole
network. 
Moreover, they adress the problem of the mismatch in a limited manner
% do not actually address the problem of topology mismatch
since they do not provide any guarantees 
that overlay and underlying topologies are
aligned with each other let alone quantify the mismatch 
and try to alleviate.
% Like all other methodologies, 
Forwarding based optimisations are commonly applied in
conjunction with other methodologies to improve 
the quality of the \p\ systems.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Caching and Replication Methodology}

%%AD in this portion there is NOT much about replication... 
%%	ALSO, references to surveyed works as missing.. (other than
%%	KAzaa!) -  Sprinkle some around the text...
Caching is widely used to exploit locality and minimize redundant
transfer of data. Caching has with much success been successfully 
adopted by web and file server application environments. 
Since peers in a \p\ system also operate as servers,
it is intuitively expected that \p\ file sharing systems can also benefit from
caching in improving performance and reducing overall resource usage. 
However, the design of caches in this context 
is non-trivial compared to the web-based caching. 
Due to the fact nodes play the double role of server and client,
two important issues have to be considered at design time. 
First, the lifetime of a query is short, as the nodes join 
and leave frequently. Second, the result
of a single query string is not always the same, as this depends on the
source of the query, the \emph{TTL} value set for the messages, 
the current interconnection of peers and the 
high volatility of the environment. 
Thus, to develop a successful caching system for a \p\ architecture, these
parameters also have to be carefully considered. 
\p\ caching/replication can be applied
at two different levels, namely caching indices or pointers to data 
(Fig.~\ref{figure:replication:index}) or caching the data itself 
(Fig.~\ref{figure:replication:data}). 
%%%
\begin{figure}[ht]
\centering
\subfigure[Indexing can reduce the cost of the last hop.] {
\includegraphics[scale=0.3]{img/pdf/replication-index.pdf}
  \label{figure:replication:index}
}\qquad\qquad
\subfigure[Data replication along the path of a successful query.] {
  \includegraphics[scale=0.3]{img/pdf/replication-data.pdf}
  \label{figure:replication:data}
}
\caption{Index and data replication strategies.}
\label{figure:replication}
\end{figure}
%%
Successful implementations have
already been developed in some commercial \p\ systems, like 
\emph{KaZaA}, which uses
caching of indices on its super-peer level. 
%%%
Even though the state of the art in \p\ protocols using caching methods
helps reduce the burden of network resources, 
their contribution in addressing the mismatch problem 
remains limited.
%%AD I am not sure what the following piece says....
%%	tried to rephrase as above...
% Unfortunately, even though the state of the art 
% in \p\ protocols using caching methods reduce the resource
% usage of the network, 
% they also fail to consistently address typical problems
% of a topologically mismatched environment such as duplication of messages.
%%%%%%%%%%%%%%%%%%%%%%%

@@

In landmark-based algorithms, nodes use network delay (e.g., \emph{RTT}) as a
distance measurement method to position themselves with respect to ``a priori''
known servers on the Internet. 
These landmark servers are used
by nodes to estimate their positions based on the intuitive assumption that
nodes
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network. 
In Figure~\ref{figure:landmarking} the newly
arriving peer, measures its distance to an array of landmark point denoted as
$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
order to choose the already participating peers with which to connect to,
compares its list to the list of the potential neighbor. Peers with similar
measured distances to the these landmark points are likely to be close to
each other as well.
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/landmarking.pdf}
\caption{Landmark binning during node bootstrap.}
\label{figure:landmarking}
\end{figure}

Landmark based protocols have four important drawbacks: First, the network
delay is not a reliable distance estimation method. For example, based on the
load on the network, the delay to certain nodes or networks can change from time
to time, which will eventually affect the distance measurements and wrong
measurements will lead to wrong estimated positions for the nodes, or incorrect
and non optimal clusterings of the nodes. Second, relying on predefined nodes
make the whole paradigm not fully distributed and the landmark system prone to
become a single point of failure. Third, using
landmark servers requires costly installation and maintenance of landmark
infrastructure across the whole Internet and for all the different autonomous
system domains. 
As popular \p\ file sharing applications usually have millions
of peers connected at any time, the network costs
of maintaining these landmark servers will likely be quite high. 
A possible solution to
the scalability problem of the static landmark servers is to use ordinary nodes
as dynamic landmarks once they estimate their own positions. Even though this
approach scales much better than static landmark servers, the measurement
accuracy problem affects the overall performance of the system. Moreover,it can
be characterized as a rather coarse-grained approximation, therefore not
particularly well suited for detecting the correct positions of nodes within
close distance to each other.
%%AD ok how about ay advantages?? Any referencess??? 
%%	Also, a bit of a discussion is missing on hybrid approaches surveyed
%%	for example which approaches use a combo of these methodologies???
